---
title: "Project291"
format: html
---

```{r}
library(readr)
library(janitor)
library(dplyr)
library(performance)
library(patchwork)
library(see)
library(ggplot2)
library(here)
library(car)
library(GGally)
```

```{r}
Life_Expectancy <- read_csv(here("Life_Expectancy_00_15.csv"))
Life<-clean_names(Life_Expectancy)
Life<-Life|>filter(year==2015)
```


```{r}
regression1<-lm(data=Life,life_expectancy~gdp_per_capita+health_expenditure+least_developed+continent+people_practicing_open_defecation+people_using_at_least_basic_drinking_water_services)
summary(regression1)

check_model(regression1, check = c("ncv","homogeneity","qq"))
```

```{r}
regression2<-lm(data=Life,life_expectancy~people_using_at_least_basic_drinking_water_services)
summary(regression2)
```

```{r}
regression3<-lm(data=Life,life_expectancy~gdp_per_capita+health_expenditure+least_developed+continent+people_practicing_open_defecation+people_using_at_least_basic_drinking_water_services+population+co2_emissions+electric_power_consumption+forest_area+individuals_using_the_internet+military_expenditure+obesity_among_adults+beer_consumption_per_capita)
summary(regression3)

anova(regression1,regression3)

regression4<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+continent+people_using_at_least_basic_drinking_water_services+co2_emissions+individuals_using_the_internet+military_expenditure+beer_consumption_per_capita)
summary(regression4)

anova(regression4,regression3)

regression5<-lm(data=Life,life_expectancy~gdp_per_capita+continent+people_using_at_least_basic_drinking_water_services+co2_emissions+individuals_using_the_internet)
summary(regression5)
anova(regression5,regression4)
```

```{r}
#regression 4 is the best model right now
regression4<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+continent+people_using_at_least_basic_drinking_water_services+co2_emissions+individuals_using_the_internet+military_expenditure+beer_consumption_per_capita)
summary(regression4)

vif(regression4)
```

$$
\begin{aligned}
\widehat{Life\_Expectancy} &= \beta_{0} + \beta_{1}(\operatorname{gdp\_per\_capita}) + \beta_{2}(\operatorname{least\_developed})\ + \\
&\quad \beta_{3}(\operatorname{continent}) + \beta_{4}(\operatorname{people\_using\_at\_least\_basic\_drinking\_water\_services}+ \\
&\quad \beta_{5}(\operatorname{co2\_emissions})+\beta_{6}(\operatorname{individuals\_using\_the\_internet}) + \\
&\quad \beta_{7}(\operatorname{military\_expenditure}) + \beta_{8}(\operatorname{beer\_consumption\_per\_capita})
\end{aligned}
$$

```{r}
regression6<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+people_using_at_least_basic_drinking_water_services+co2_emissions+individuals_using_the_internet+military_expenditure+beer_consumption_per_capita)
summary(regression6)
```

```{r}
regression7<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+people_using_at_least_basic_drinking_water_services+co2_emissions+individuals_using_the_internet)
summary(regression7)
anova(regression7,regression4)
#regression 4 is better
```

```{r}
regression8<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+people_using_at_least_basic_drinking_water_services+co2_emissions+military_expenditure+beer_consumption_per_capita)
summary(regression8)
```


```{r}
Life_numeric<-Life|>
  select(-continent,-least_developed,-country,-year)
GGally::ggpairs(Life_numeric)
```

```{r}
regression9<-lm(data=Life,life_expectancy~gdp_per_capita+least_developed+people_using_at_least_basic_drinking_water_services+co2_emissions+health_expenditure+people_practicing_open_defecation)
summary(regression9)
check_model(regression9, check = c("ncv","homogeneity","qq"))
vif(regression9)
#regression 4 or regression9 (not nested, larger R-square for regression4, VIF better shown in regression9)
```

$$
\begin{aligned}
\widehat{Life\_Expectancy} &= \beta_{0} + \beta_{1}(\operatorname{gdp\_per\_capita}) + \beta_{2}(\operatorname{least\_developed})\ + \\
&\quad \beta_{3}(\operatorname{people\_using\_at\_least\_basic\_drinking\_water\_services}+ \\
&\quad \beta_{4}(\operatorname{co2\_emissions})+\beta_{5}(\operatorname{beer\_consumption\_per\_capita}) + \\
&\quad \beta_{6}(\operatorname{health\_expenditure}) + \beta_{7}(\operatorname{people\_practicing\_open\_defecation})
\end{aligned}
$$
# additive models:

```{r}
regression_a <- lm(data = Life, life_expectancy ~ people_using_at_least_basic_drinking_water_services + individuals_using_the_internet + gdp_per_capita + obesity_among_adults + people_practicing_open_defecation + electric_power_consumption + health_expenditure + co2_emissions + beer_consumption_per_capita + military_expenditure + population + forest_area + continent + least_developed)
summary(regression_a)
```

Based on p-value in the regression table for each variables, it seems that obesity_among_adults, electric_power_consumption, health_expenditure,  beer_consumption_per_capita, population, and forest_area are not statistically significant to the regression model.  

```{r}
regression_b <- lm(data = Life, life_expectancy ~ people_using_at_least_basic_drinking_water_services + individuals_using_the_internet + gdp_per_capita + people_practicing_open_defecation + co2_emissions + military_expenditure + continent + least_developed)
summary(regression_b)
vif(regression_b)
```

p-value qualified, vif in danger: people_using_at_least_basic_drinking_water_services(5.3729), individuals_using_the_internet(7.6596)

```{r}
Life |>
  select(people_using_at_least_basic_drinking_water_services, individuals_using_the_internet, gdp_per_capita, people_practicing_open_defecation, co2_emissions, military_expenditure, continent, least_developed) |>
  ggpairs()
```

high correlation value: gdp_per_capita + individuals_using_the_internet (0.811), etc.

Overall, we try to delete individuals_using_the_internet.

```{r}
regression_c <- lm(data = Life, life_expectancy ~ people_using_at_least_basic_drinking_water_services + gdp_per_capita + people_practicing_open_defecation + co2_emissions + military_expenditure + continent + least_developed)
summary(regression_c)
vif(regression_c)
```

p value for least_develope = 0.0654, removed
VIFs are generally reasonable

```{r}
regression_d <- lm(data = Life, life_expectancy ~ people_using_at_least_basic_drinking_water_services + gdp_per_capita + people_practicing_open_defecation + co2_emissions + military_expenditure + continent)
library(equatiomatic)
extract_eq(regression_d)

summary(regression_d)
vif(regression_d)
```
$$
\operatorname{life\_expectancy} = \alpha + \beta_{1}(\operatorname{people\_using\_at\_least\_basic\_drinking\_water\_services})+ \beta_{2}(\operatorname{gdp\_per\_capita}) + \beta_{3}(\operatorname{people\_practicing\_open\_defecation}) + \beta_{4}(\operatorname{co2\_emissions}) + \beta_{5}(\operatorname{military\_\expenditure}) + \beta_{6}(\operatorname{continent}_{\operatorname{Asia}}) + \beta_{7}(\operatorname{continent}_{\operatorname{Europe}}) + \beta_{8}(\operatorname{continent}_{\operatorname{North\ America}}) + \beta_{9}(\operatorname{continent}_{\operatorname{Oceania}}) + \beta_{10}(\operatorname{continent}_{\operatorname{South\ America}}) + \epsilon
$$
approved.

```{r}
plot_check <- check_model(regression_d, check=c("linearity", "qq"))
plot(plot_check, data=regression_d)
```

consumption checked

# interaction models:

```{r}
Life |>
  select(people_using_at_least_basic_drinking_water_services, gdp_per_capita, people_practicing_open_defecation, co2_emissions, military_expenditure, continent) |>
  ggpairs()
```

strong relationship between: people_practicing_open_defecation and people_using_at_least_basic_drinking_water_services (-0.748), co2_emissions and gdp_per_capita (0.770), gdp_per_capita and people_using_at_least_basic_drinking_water_services (0.559)

```{r}
regression_e <- lm(data = Life, life_expectancy ~ people_using_at_least_basic_drinking_water_services * people_practicing_open_defecation + gdp_per_capita * co2_emissions + military_expenditure + continent)
summary(regression_e)
anova(regression_d, regression_e)
```

```{r}
plot_check <- check_model(regression_e, check=c("linearity", "qq"))
plot(plot_check, data=regression_e)
```

```{r}
reg_continent<-lm(data=Life, life_Expectancy~continent)
summary(reg_continent)
```


